---
title: "Validation evidence: P. vivax"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Validation evidence: P. vivax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
claims <- fleetcheck::read_claims(
  system.file("claims.yml", package = "fleetcheck"))
if (!nrow(claims)) claims <- fleetcheck::read_claims("../claims.yml")
claims <- claims[claims$parasite == "vivax", ]
```

How closely does `fleet` reproduce `malariasimulation` for *P. vivax*? The same
questions as [for *P. falciparum*](evidence.html), for a
`get_parameters(parasite = "vivax")` list: on vivax's own transmission grid, EIR
0.3 to 30, and with radical cure, by primaquine and by tafenoquine, in place of
the chemoprevention and vaccines `malariasimulation` cannot run under vivax.
Vivax has no severe disease, so there are no severe claims, and the site-file
comparison of tier 3 is not yet part of the vivax evidence.

One thing bears on several verdicts below. `malariasimulation` 3.0.0 evaluates
vivax relapses only on days when at least one person is bitten, so wherever
days without a bite are common -- under strong vector control, at low
transmission -- the IBM loses relapses that `fleet` keeps. `intervention-impact-pv`
sets out what that does to the comparison.

Every section is generated from `claims.yml`, so this page cannot disagree with
the register, and the register cannot be changed without the page following.

```{r scoreboard, results = "asis"}
# link_prefix = "" links each id to its own section on this page
cat(fleetcheck::scoreboard_md(claims, link_prefix = ""), sep = "\n")
```

```{r sections, results = "asis"}
cat(fleetcheck::claim_sections_md(claims), sep = "\n")
```

## What "tier" means

The same as [for *P. falciparum*](evidence.html#what-tier-means): what a result
costs to reproduce, recorded against each claim. Every vivax claim here is tier
2, re-running both models over the vivax scenario suite: about an hour on ten
cores.
