---
title: "Validation evidence"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Validation evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
claims <- fleetcheck::read_claims(
  system.file("claims.yml", package = "fleetcheck"))
if (!nrow(claims)) claims <- fleetcheck::read_claims("../claims.yml")
```

How closely does `fleet` reproduce `malariasimulation`? Each claim below states
the criterion that decides it, the value measured against that criterion, and a
verdict. The order runs from transmission through clinical and severe burden to
their age distributions, then interventions and real settings, with the
demographic, performance and numerical checks that underwrite them at the end.

Every section is generated from `claims.yml`, so this page cannot disagree with
the register, and the register cannot be changed without the page following.

```{r scoreboard, results = "asis"}
# link_prefix = "" links each id to its own section on this page
cat(fleetcheck::scoreboard_md(claims, link_prefix = ""), sep = "\n")
```

```{r sections, results = "asis"}
for (i in seq_len(nrow(claims))) {   # register order, same as the table above
  cl <- claims[i, ]
  cat(sprintf("\n## %s {#%s}\n\n", cl$id, cl$id))
  # verdict_html(), not a second copy of the map: this file used to hold one in
  # different case, so a claim read `pass` in the table and `PASS` here
  cat(sprintf("%s &mdash; %s\n\n", fleetcheck::verdict_html(cl$status), cl$claim))
  # The criterion and the measurement are in the table above and are not
  # repeated here. What this line carries is what a reader needs while looking
  # at the figure: what the result cost to produce, and where the code that
  # made it is.
  cat(sprintf('<p class="claim-meta">tier %d &middot; <code>%s</code></p>\n\n',
              cl$tier, cl$evidence))
  if (nzchar(cl$note)) cat(cl$note, "\n\n")

  # Each claim shows the figure that is evidence for THAT claim, and nothing
  # else: a figure showing four things is not evidence for any one of them.
  fig <- cl$figure
  if (nzchar(fig) && !fig %in% c("NA", "~") && file.exists(fig)) {
    # Alt text is the claim the figure is evidence for. An empty alt would tell
    # a screen reader the image is decorative, on a page whose entire content is
    # evidence.
    # Written as an <img> rather than as `![alt](src)`, because pandoc turns a
    # markdown image into a <figure> and prints the alt text as a caption --
    # which put the claim on screen a second time, directly under the heading
    # that had just said it. Every plot carries its own title and footnote.
    # Width and height are emitted so the browser can reserve the space before
    # the image arrives. Without them the figures are 29% of the page height
    # with nothing holding it open, so a deep link to a claim anchor -- which is
    # how every link from the README and the front page arrives -- jumps to a
    # position that then moves hundreds of pixels as the images load.
    # max-width: 100% in the stylesheet still scales them down.
    d <- tryCatch(dim(png::readPNG(fig)), error = function(e) NULL)
    cat(sprintf('<img src="%s"%s alt="%s">\n\n', fig,
                if (length(d) >= 2) sprintf(' width="%d" height="%d"', d[2], d[1]) else "",
                gsub('"', "&quot;", cl$claim, fixed = TRUE)))
  } else {
    cat("*No figure: this claim is a single number rather than a relationship",
        "across a range, and a chart of one number carries less than the",
        "number.*\n\n")
  }
}
```

## What "tier" means

What it costs to reproduce a result is a property of that result, so it is
recorded against every claim rather than stated once in prose.

| tier | cost | who can reproduce it |
| --- | --- | --- |
| 0 | seconds | anyone, from the committed summaries |
| 1 | ~2 min | anyone; re-runs `fleet` only |
| 2 | ~2 h | anyone with about ten cores |
| 3 | ~40 min | about ten cores, and inputs that are not redistributable |

Tier 3 is the 63-country site-file comparison. **Its figures and statistics are
public; the site files behind them are not.** The constraint is the inputs
rather than the compute: the sweep re-runs `fleet` only, because the IBM arm is
the pre-run diagnostic shipped with each site file, so forty minutes on ten
cores refreshes it. But the site files are not redistributable, so nobody
outside the project can repeat it. That is a limitation of this evidence, not a
property of the result, and
`validations/03-real-settings/example-one-site.R` runs the same pipeline on a
single sub-site for anyone who has one site file.
