---
title: "The evidence, claim by claim"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{The evidence, claim by claim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
claims <- fleetcheck::read_claims(
  system.file("claims.yml", package = "fleetcheck"))
if (!nrow(claims)) claims <- fleetcheck::read_claims("../claims.yml")
```

Every section below is generated from `claims.yml`. Nothing on this page is typed
by hand, so it cannot disagree with the register, and the register cannot be
edited without this page following.

```{r scoreboard, results = "asis"}
# link_prefix = "" targets anchors on this page: the scoreboard at the top of
# the evidence article used to be the one copy with no links at all, so the
# reader met eleven verdicts and then had to scroll to find any of them.
cat(fleetcheck::scoreboard_md(claims, link_prefix = ""), sep = "\n")
```

```{r sections, results = "asis"}
label <- c(pass = '<span class="verdict pass">PASS</span>',
           open = '<span class="verdict open">OPEN</span>',
           fail = '<span class="verdict fail">FAIL</span>',
           undeclared = '<span class="verdict untested">UNTESTED</span>')
ord <- order(match(claims$status, c("fail", "undeclared", "open", "pass")),
             claims$tier)

for (i in ord) {
  cl <- claims[i, ]
  cat(sprintf("\n## %s {#%s}\n\n", cl$id, cl$id))
  cat(sprintf("%s &mdash; %s\n\n", label[[cl$status]], cl$claim))
  cat("| | |\n| --- | --- |\n")
  cat(sprintf("| Criterion | %s |\n", fleetcheck::md_cell(cl$criterion)))
  cat(sprintf("| Measured | %s |\n", fleetcheck::md_cell(cl$measured)))
  cat(sprintf("| Tier | %d |\n", cl$tier))
  cat(sprintf("| Evidence | `%s` |\n\n", cl$evidence))
  if (nzchar(cl$note)) cat(cl$note, "\n\n")

  # Each claim shows the figure that is evidence for THAT claim. The four
  # transmission-intensity claims each get their own panel rather than sharing
  # the four-panel composite, because a figure showing four things is not
  # evidence for any one of them.
  fig <- cl$figure
  if (nzchar(fig) && !fig %in% c("NA", "~") && file.exists(fig)) {
    # Alt text is the claim the figure is evidence for. Every figure here was
    # shipped with an empty `![]()`, which tells a screen reader the image is
    # decorative -- on a page whose entire content is evidence.
    cat(sprintf("![%s](%s)\n\n", cl$claim, fig))
  } else {
    cat("*No figure. This claim is a single measured quantity rather than a",
        "relationship across a range, and a chart of one number carries less",
        "than the number.*\n\n")
  }
}
```

## Systematic bias: how far apart the two models sit

The claims above ask whether `fleet` **tracks** the IBM — whether it has the
right shape, the right direction, and lands inside the IBM's own stochastic
spread. They do not ask whether it sits exactly on top of it, and it does not.

That is expected rather than a defect, and it is not a pass-or-fail question, so
it is quantified here instead of being carried as a claim with a verdict. A
mean-field model and an individual-based one differ for a structural reason.
`fleet` evaluates each immunity-dependent probability **once per stratum, at
that stratum's mean immunity**; the IBM averages the same function over
individuals whose exposure histories differ *within* the stratum. Averaging a
convex function the second way gives the larger answer, so the mean field
returns the smaller one. Severe disease is the most exposed to this because
$I_{V0}$ is sixteen times smaller than $I_{C0}$, which puts $\theta$ out on its
convex tail by age 3 while $\phi$ is still nearly linear.

**The size of that bias is not yet settled, and the register should not pretend
otherwise.** Two things are in the way.

### It is entangled with the age grid

`fleet` represents age as 52 discrete groups, and within a group everyone
carries one immunity value. That is a first-order approximation, and for severe
incidence it is not converged. Refining the grid — each level subdividing every
band of the default grid, so the comparison-band edges survive — moves all-age
severe incidence against the IBM median like this:

| age groups | EIR 3 | EIR 20 | EIR 120 |
| --- | --- | --- | --- |
| 52 (the default) | −2.7% | −2.8% | −3.5% |
| 103 | −0.6% | +0.6% | +1.7% |
| 205 | +0.5% | +2.4% | +4.5% |
| 409 | **+1.1%** | **+3.3% ** | **+6.0%** |

The convergence is clean and first-order, and the values are the model's rather
than the solver's — they hold to 1e-13 as the tolerance is tightened from 1e-8
to 1e-11. **So the deficit visible on the default grid has the opposite sign to
the converged answer.** Prevalence is converged and does not move; clinical
incidence behaves like severe but less so.

What that means for a reader: the mean-field argument above sets the direction,
and a discretisation error of comparable size and opposite sign sits on top of
it. Quoting either number alone would be misleading. Intervention **impact** —
the reduction one deployment produces, which is what `fleet` is mostly used for
— is far less affected, because it is a ratio of two runs that share the grid:
a four-fold refinement moves it by at most 0.9 percentage points.

### Across real settings it is a range, not a number

The 63-country comparison reports `fleet` about 9% above the IBM on clinical and
severe incidence across 1,391 sub-sites. That is a mean over a strong structure,
and the mean on its own is the least useful summary of it. Stratified by the
IBM's own incidence and by treatment coverage, the ratio of `fleet` to the IBM
on clinical incidence runs:

| | lowest coverage | → | highest coverage |
| --- | --- | --- | --- |
| highest transmission | **1.00** | 1.03 | 1.07 |
| lowest transmission | 1.16 | 1.54 | 1.70 |

At high transmission with low treatment coverage the two models agree almost
exactly. The excess grows as transmission falls and as treatment coverage rises,
and within Burkina Faso alone it runs from 0.96 to 1.08. A single figure of
"+9%" describes none of those cells well.

A controlled test of the treatment half of that — one setting, one EIR, coverage
swept from 0 to 0.8 against 20 IBM replicates — found the effect on incidence
statistically real but **smaller than the IBM's own replicate spread**, so most
of the observational gradient is other things varying alongside coverage. The
site-file numbers also predate a `fleet` change that affects exactly the
time-varying demography every site file carries.

### What to take from this

- `fleet` tracks the IBM's shape well: prevalence agrees to about 1% everywhere
  tested, and correlation across 1,391 real settings is 0.98 on clinical.
- On **levels**, expect a few per cent, in a direction and of a size that depends
  on the age grid and on the setting. Do not quote a single bias figure.
- On **intervention impact**, the agreement is much tighter than either of the
  above, and that is the use this matters least for.

## What "tier" means

Reproducing a result costs something, and that cost is a property of the result,
so it is recorded against every claim rather than mentioned once in prose.

| tier | cost | who can reproduce it |
| --- | --- | --- |
| 0 | seconds | anyone, from the committed summaries |
| 1 | ~2 min | anyone; re-runs `fleet` only |
| 2 | ~25 min | anyone with about ten cores |
| 3 | ~7 h | a cluster, and inputs that are not redistributable |

Tier 3 is the 63-country site-file comparison. **Its figures and statistics are
public; the underlying runs are not.** The code that produced them is in this
repository and can be read and audited. Re-running it needs the malariaverse
site files and a cluster, so nobody outside the project can repeat it, and that
is a limitation of this evidence rather than a property of the result.
