---
title: "Validation evidence: P. vivax"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Validation evidence: P. vivax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
# the source register when built from the source tree (see methods.Rmd)
claims <- fleetcheck::read_claims(if (file.exists("../claims.yml")) "../claims.yml" else
  system.file("claims.yml", package = "fleetcheck"))
claims <- claims[claims$parasite == "vivax", ]
```

How closely does `fleet` reproduce `malariasimulation` for *P. vivax*? The same
questions as [for *P. falciparum*](evidence.html), for a
`get_parameters(parasite = "vivax")` list: on vivax's own transmission grid, EIR
0.3 to 30, with radical cure, by primaquine and by tafenoquine, in place of the
chemoprevention `malariasimulation` cannot run under vivax and the vaccine it has
no vivax calibration for. `malariasimulation`'s vivax model has no severe
disease, so there are no severe claims. Two claims are vivax's own: relapse
incidence and the share of people carrying hypnozoites. The site-file comparison
of tier 3 is not yet part of the vivax evidence.

One thing bears on the verdicts below. `malariasimulation` 3.0.0 evaluates vivax
relapses only on days when at least one person is bitten, so on a day with no
infectious bite anywhere the IBM loses every relapse, which `fleet` keeps. In the
scenario suite's 10,000-person runs such days are common under indoor residual
spraying, and below an EIR of about 0.14 whatever is deployed;
`intervention-impact-pv` sets out what that does to the comparison.

Every section is generated from `claims.yml`, so this page cannot disagree with
the register, and the register cannot be changed without the page following.

```{r scoreboard, results = "asis"}
# link_prefix = "" links each id to its own section on this page
cat(fleetcheck::scoreboard_md(claims, link_prefix = ""), sep = "\n")
```

```{r sections, results = "asis"}
cat(fleetcheck::claim_sections_md(claims), sep = "\n")
```

## What "tier" means

The same as [for *P. falciparum*](evidence.html#what-tier-means): what a result
costs to reproduce, recorded against each claim. The costs are vivax's own.

| tier | cost | who can reproduce it |
| --- | --- | --- |
| 1 | a few minutes | anyone; re-runs `fleet` only |
| 2 | about an hour | anyone with about ten cores |
