---
title: "The evidence, claim by claim"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{The evidence, claim by claim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
claims <- fleetcheck::read_claims(
  system.file("claims.yml", package = "fleetcheck"))
if (!nrow(claims)) claims <- fleetcheck::read_claims("../claims.yml")
```

Every section below is generated from `claims.yml`. Nothing on this page is typed
by hand, so it cannot disagree with the register, and the register cannot be
edited without this page following.

```{r scoreboard, results = "asis"}
# link_prefix = "" targets anchors on this page: the scoreboard at the top of
# the evidence article used to be the one copy with no links at all, so the
# reader met eleven verdicts and then had to scroll to find any of them.
cat(fleetcheck::scoreboard_md(claims, link_prefix = ""), sep = "\n")
```

```{r sections, results = "asis"}
label <- c(pass = "PASS", open = "OPEN", fail = "FAIL",
           undeclared = "UNTESTED")
declared <- c(`in-advance` = "declared before the run",
              retrospective = "written retrospectively, at first pass",
              none = "no criterion exists")
ord <- order(match(claims$status, c("fail", "undeclared", "open", "pass")),
             claims$tier)

for (i in ord) {
  cl <- claims[i, ]
  cat(sprintf("\n## %s {#%s}\n\n", cl$id, cl$id))
  cat(sprintf("**%s** &mdash; %s\n\n", label[[cl$status]], cl$claim))
  cat("| | |\n| --- | --- |\n")
  cat(sprintf("| Criterion | %s |\n", fleetcheck::md_cell(cl$criterion)))
  cat(sprintf("| Measured | %s |\n", fleetcheck::md_cell(cl$measured)))
  cat(sprintf("| Tier | %d |\n", cl$tier))
  cat(sprintf("| Evidence | `%s` |\n", cl$evidence))
  cat(sprintf("| Criterion was | %s |\n\n", declared[[cl$declared]]))
  if (nzchar(cl$note)) cat(cl$note, "\n\n")

  # Each claim shows the figure that is evidence for THAT claim. The four
  # transmission-intensity claims each get their own panel rather than sharing
  # the four-panel composite, because a figure showing four things is not
  # evidence for any one of them.
  fig <- cl$figure
  if (nzchar(fig) && !fig %in% c("NA", "~") && file.exists(fig)) {
    # Alt text is the claim the figure is evidence for. Every figure here was
    # shipped with an empty `![]()`, which tells a screen reader the image is
    # decorative -- on a page whose entire content is evidence.
    cat(sprintf("![%s](%s)\n\n", cl$claim, fig))
  } else {
    cat("*No figure. This claim is a single measured quantity rather than a",
        "relationship across a range, and a chart of one number carries less",
        "than the number.*\n\n")
  }
}
```

## What "tier" means

Reproducing a result costs something, and that cost is a property of the result,
so it is recorded against every claim rather than mentioned once in prose.

| tier | cost | who can reproduce it |
| --- | --- | --- |
| 0 | seconds | anyone, from the committed summaries |
| 1 | ~2 min | anyone; re-runs `fleet` only |
| 2 | ~25 min | anyone with about ten cores |
| 3 | ~7 h | a cluster, and inputs that are not redistributable |

Tier 3 is the 63-country site-file comparison. **Its figures and statistics are
public; the underlying runs are not.** The code that produced them is in this
repository and can be read and audited. Re-running it needs the malariaverse
site files and a cluster, so nobody outside the project can repeat it, and that
is a limitation of this evidence rather than a property of the result.
