---
title: "Validation evidence: P. falciparum"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Validation evidence: P. falciparum}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
# the source register when built from the source tree (see methods.Rmd)
claims <- fleetcheck::read_claims(if (file.exists("../claims.yml")) "../claims.yml" else
  system.file("claims.yml", package = "fleetcheck"))
claims <- claims[claims$parasite == "falciparum", ]
```

How closely does `fleet` reproduce `malariasimulation` for *P. falciparum*? Each
claim below states the criterion that decides it, the value measured against
that criterion, and a verdict. The order runs from transmission through clinical
and severe burden to their age distributions, then interventions and real
settings, with the demographic, performance and numerical checks that underwrite
them at the end. The *P. vivax* claims are on [a page of their
own](evidence-vivax.html).

Every section is generated from `claims.yml`, so this page cannot disagree with
the register, and the register cannot be changed without the page following.

```{r scoreboard, results = "asis"}
# link_prefix = "" links each id to its own section on this page
cat(fleetcheck::scoreboard_md(claims, link_prefix = ""), sep = "\n")
```

```{r sections, results = "asis"}
cat(fleetcheck::claim_sections_md(claims), sep = "\n")
```

## What "tier" means

What it costs to reproduce a result is a property of that result, so it is
recorded against every claim rather than stated once in prose.

| tier | cost | who can reproduce it |
| --- | --- | --- |
| 0 | seconds | anyone, from the committed summaries |
| 1 | minutes | anyone; re-runs `fleet` only |
| 2 | ~2 h | anyone with about ten cores |
| 3 | ~20 min | four cores, and inputs that are not redistributable |

Tier 3 is the 63-country site-file comparison. **Its figures and statistics are
public; the site files behind them are not.** The constraint is the inputs
rather than the compute: the sweep re-runs `fleet` only, because the IBM arm is
the pre-run diagnostic shipped with each site file, so twenty minutes on four
cores refreshes it. But the site files are not redistributable, so nobody
outside the project can repeat it. That is a limitation of this evidence, not a
property of the result, and
`validations/03-real-settings/example-one-site.R` runs the same pipeline on a
single sub-site for anyone who has one site file.
